---
title: "birdsize"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{birdsize}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(birdsize)
```

The core functions in `birdsize` apply at 3 levels of organization: species, population, and community.

The `species_*` functions take information about a real or hypothetical species and generates the parameters necessary to simulate body size distributions for that species.

The `population_*` functions use species-level parameters and abundances (population sizes) to simulate individual body size and basal metabolic rate measurements to make up populations of that species, and calculate population-wide summary statistics. 

The `community_*` functions generate population-level estimates for numerous populations (e.g. populations of different species, or populations of the same species at different points in time or different sampling locations), and calculate summary statistics with flexible grouping structures.

This vignette walks through each of these operations in detail.

## Species-level parameters

## Population-level estimates

## Community-level simulations
