---
title: "Simulate populations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sim_populations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(22)
```

```{r setup, echo = F}
library(birdsize)
library(ggplot2)
theme_set(theme_bw())
```

# Simulate populations of individuals of a single species

The core functionality of `birdsize` is to simulate individual body mass measurements for birds given either their species ID or a mean (and possibly standard deviation) of body mass. `pop_generate` achieves this for a single species. `comm_generate` uses `pop_generate` to simulate body masses for multiple populations of birds of different species. 


## Using species identity

For most birds in the North American Breeding Bird Survey, you can use the species' code (AOU) to simulate a population directly. For the hummingbird _Selasphorus calliope_: 

```{r, fig.dim =c(3.5,3)}

a_hundred_hummingbirds <- pop_generate(abundance = 100, aou = 244)

head(a_hundred_hummingbirds)

ggplot(a_hundred_hummingbirds, aes(individual_mass)) + 
  geom_histogram(bins = 25) +
  xlab("Mass (g)") +
  ylab("Count") +
  ggtitle("A population of hummingbirds") 

```

To look up the AOU based on the species' scientific name, explore the included `sd_table` dataset:

```{r}
sd_table

```

## Using a known mean and standard deviation

Alternatively, you  can simulate body masses for a population by supplying the body size parameters yourself. This may be useful if you would like to work with a species not included in the `sd_table`, test sensitivities to different parameter ranges, or generate values for simulation/null models (or, other applications!). 

**Note that, if both mean mass and a species code are provided, the species code will be used and the mean mass provided will be ignored!**

```{r, fig.dim =c(3.5,3)}

a_hundred_hypotheticals <- pop_generate(abundance = 100, mean_size = 25, sd_size = 3)

head(a_hundred_hypotheticals)

ggplot(a_hundred_hypotheticals, aes(individual_mass)) + 
  geom_histogram(bins = 25) +
  xlab("Mass (g)") +
  ylab("Count") +
  ggtitle("A population of hypothetical birds", subtitle ="Mean mass = 25 g\nStandard deviation = 3") 


```

## Using a known mean, but no standard deviation

If the mean mass is not known or not provided, `simulate_population` will estimate the standard deviation based on scaling between the mean and standard deviation of body mass:


```{r, fig.dim =c(3.5,3)}


another_hundred_hypotheticals <- pop_generate(abundance = 100, mean_size = 25)

head(another_hundred_hypotheticals)

ggplot(another_hundred_hypotheticals, aes(individual_mass)) + 
  geom_histogram(bins = 25) +
  xlab("Mass (g)") +
  ylab("Count") +
  ggtitle("A population of hypothetical birds", subtitle ="Mean mass = 25 g\nStandard deviation = 1.74") 


```
