---
title: "species"
output: 
  rmarkdown::html_vignette:
      toc: true
vignette: >
  %\VignetteIndexEntry{species}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(22)
```

```{r setup, echo = F}
library(birdsize)
library(ggplot2)
library(purrr)
theme_set(theme_bw())
```

The `species_define` function works mostly behind the scenes to set up the parameters needed to simulate individual body size measurements. Given any of: (1) the AOU as used in the North American Breeding Bird Survey (Pardieck et al. 2019), (2) the scientific name, (3) the species' mean body size, or (4) the species' mean _and standard deviation_ body size, `species_define` returns the parameters used by `pop_generate` and `community_generate` to simulate individual size measurements for individuals of that species, or returns an error message asking for more or different information. In most instances, `species_define` is called under-the-hood by the `generate` functions, and users do not need to interact with it directly. 

# Species known to `birdsize`

`birdsize` includes species-level parameters for 443 species of birds common in the North American Breeding Bird Survey. To view the list of species included, examine the `known_species` data table, included:

```{r}

head(known_species)

```

Species included in `known_species` can be retrieved via either their AOU or scientific name. For example, the hummingbird _Selasphorous calliope_ has an AOU of `4360`. 

## AOU lookup

```{r}

hummingbird_AOU_parameters <- species_define(AOU = 4360)

hummingbird_AOU_parameters

```

## Scientific name lookup

```{r}

hummingbird_name_parameters <- species_define(scientific_name = "Selasphorus calliope")

hummingbird_name_parameters

```

Note that the `sd_method` field tells us which method we used to look up the parameters. This field propagates throughout the `pop_generate` and `community_generate` functions to keep track of the underlying methodology. 

## Unknown species or AOUs

Attempting to use `species_define` with an AOU or species not known to `birdsize` will return an error:

```{r}

try(species_define(AOU = 100))

```


```{r}

try(species_define(scientific_name = "Taylor swift"))

```

# Species not known to `birdsize`

Some users may want to use this methodology with species not included in `known_species`, or to use different species-level parameters than those built-in to `birdsize` (for example, to explore intraspecific variation in body size over time or space). To do this, supply `species_define` with mean, or mean and standard deviation, values for each species. To help keep track of species-parameter matches, use the `sim_species_id` field to assign a species identifier to each novel species. 

## Manually supplying species parameters

Suppose we want to work with a hypothetical species with a mean body size of 40g and a standard deviation of 2.5. Because this species doesn't have a scientific name or AOU included in `birdsize`, we label it using the arbitrary `sim_species_id` of 1.

```{r}

hypothetical_species_parameters <- species_define(mean_size = 40, sd_size = 2.5, sim_species_id = 1)

hypothetical_species_parameters

```

This can be particularly useful when working with multiple new species. For example, if we have 3 new species, we can store their information in a separate table and iterate over `sim_species_id` to generate parameters for each species. This happens under the hood in `community_generate`.

```{r}

multiple_species_info <- data.frame(
  mean_size = c(10, 40, 50),
  sd_size = c(1, 2.5, 3),
  sim_species_id = 1:3
) 


pmap_df(multiple_species_info, species_define)

```

If the standard deviation is not provided, `species_define` will estimate it (see the `scaling` vignette):

```{r}


multiple_species_info_no_sd <- data.frame(
  mean_size = c(10, 40, 50),
  sim_species_id = 1:3
) 


pmap_df(multiple_species_info_no_sd, species_define)

```

# Order of operations

If multiple sets of information are provided (e.g. both `AOU` and `scientific_name`), `species_define` will use it in this order of preference: 

1. AOU
2. Scientific name
3. Manually provided mean and standard deviation
4. Manually provided mean and estimated standard deviation


# References


Pardieck, K.L., Ziolkowski Jr., D.J., Lutmerding, M., Aponte, V., and Hudson, M-A.R., 2019, North American Breeding Bird Survey Dataset 1966 - 2018 (ver. 2018.0): U.S. Geological Survey, Patuxent Wildlife Research Center, https://doi.org/10.5066/P9HE8XYJ.
